%   Dylan Wright - dylan.wright@uky.edu
%   EE480 - Assignment 2: The Making Of An IDIOT
%   note.tex : Implementer's Note
%   Version:
%       02-14-2016 : initial

\documentclass[conference]{IEEEtran}
\usepackage{graphicx}

\begin{document}
\title{Assignment 2: The Making Of An IDIOT\\Implementor's Notes}
\author{\IEEEauthorblockN{Dylan Wright}
        \IEEEauthorblockA{dylan.wright@uky.edu}
        \IEEEauthorblockN{Casey O'Kane}}

\maketitle

\begin{abstract}
\end{abstract}

\section{Testing}
\subsection{Instruction Set Architecture}
In order to test the IDIOT instruction set specification a test framework was
implemented. This framework is in the \texttt{IDIOT/} directory. The framework
consists of the folowing files:
\subsubsection{\texttt{aik.py}}
To automatically test files \texttt{aik.py} sends a PUSH request to the
AIK cgi program. The returned html page is parsed and each section is output.
The \texttt{.text} and \texttt{.data} sections are sent to stdout and the
assembler messages are sent to \texttt{stderr}. This method is not ideal, an
AIK executable would be prefferable. Sample run:
\begin{verbatim}
 $ echo "file.idiot" | ./aik.py
\end{verbatim}
\subsubsection{\texttt{diss.py}}
To make test results human readable, \texttt{diss.py} dissassembles a 
\texttt{.out} file (the \texttt{.text} and \texttt{.data} segment of the 
output of \texttt{aik.py}). The code is converted to binary and displayed in
tabular format. Sample run:
\begin{verbatim}
 $ echo "file.out" | ./diss.py
\end{verbatim}
\subsubsection{\texttt{test.sh}}
This file can be used to test each \texttt{.idiot} file in the \texttt{progs/}
directory. This script runs each file through AIK and compares the output to
the expected output. \texttt{.text} and \texttt{.data} segment expected
output should be placed in a file with the same name as the program and a 
\texttt{.expected.out} file extension. Expected assembler messages should
be placed in a file with a \texttt{.expected.err} extension. The test script
will report the number of passed, failed and possibly failed tests. This test
framework was adapted from a script provided by Dr. Jaromczyk in the Fall
2015 CS 441G: Compilers course. Sample run:
\begin{verbatim}
 $ ./test
\end{verbatim}
\subsection{Verilog Modules}
\end{document}
